---
title: "A comparison between evenness(PIE) change and Species Richness/Assemblage Size change"
author: "Faith Jones"
date: "5 June 2017"
output: html_document
---
```{r, include=FALSE}
library(ggplot2)
library(ggExtra)
library(vegan)
library(reshape2)
```


```{r, include=FALSE}
evennessSim <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/evennessChangeSimp.csv")
evenness <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/evennessChange.csv")
Abundance <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/AbundanceCapacityChange.csv")
AbDominance <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/ActualAbundanceChange.csv")
RelDominance <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/PercentAbundanceChange.csv")
rareity <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/rarityChange.csv")
rareity2 <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/rarityChangeSingelton.csv")
SRChange <- read.csv("C:/Users/faj/Documents/OneDrive for Business/research/ch3 - metaanalysis/data/SpeciesRichnessChange.csv")

##renaming columns so they make sense 
names(Abundance)[4] <- "AbundanceAll"
names(evennessSim)[4] <- "EvennessSim1"
names(evenness)[4] <- "Evenness"
names(AbDominance)[4] <- "AbundanceDom"
names(RelDominance)[4] <- "relativeDom"
names(rareity)[4] <- "Rareity"
names(rareity2)[4] <- "Rareity2"

##merging all data
AbEv <- merge(Abundance, evenness, by = "Study")
AbRel <- merge(AbDominance, RelDominance, by = "Study")
RareChange <- merge(rareity,AbRel, by = "Study")
RareChange2 <- merge(rareity2,RareChange, by = "Study")
Allchange1 <- merge(AbEv,RareChange2, by = "Study")
AllSR <- merge(SRChange,Allchange1, by.x = "Study_ID", by.y = "Study")
Allchange <- merge(evennessSim, AllSR,  by.y = "Study_ID", by.x = "Study")
names(Allchange)

#selecting only columns with coefficients of change, and not all teh
#extra info like taxa or ;ength of study
selectData <- Allchange[,c(1,4, 13, 18, 29, 38, 44, 50, 60, 5, 61, 62, 41)]
head(selectData)

##plot scatter plot with marginal histograms
#evennes-assemblage size
evAss <- ggplot(selectData, aes( AbundanceAll, EvennessSim1)) + 
	geom_point() + 
	theme_classic()
ggMarginal(evAss, type = "histogram")

#chaneg taxa to just invert/vert/plant
levels(selectData$TAXA.x)
levels(selectData$TAXA.x)[1] <- "Multipe"
levels(selectData$TAXA.x)[2] <- "Vertebrate"
levels(selectData$TAXA.x)[4] <- "Vertebrate"
levels(selectData$TAXA.x)[4] <- "Vertebrate"
levels(selectData$TAXA.x)[5] <- "Vertebrate"
levels(selectData$TAXA.x)[6] <- "Vertebrate"
levels(selectData$TAXA.x)[3] <- "Multipe"
selectData <- unique(selectData)

```

##Description of the plots
The below plots are to see if there are any interesting patterns in the 
relationship between evenness change and assemblage size or species richness. 
Before plotting I checked how many assemblages are in each type of taxa, realm
and protected area stuatus. The high proportion of plants in protected areas
makes it diffiult to interpret the affects of protected area on change. 

```{r}
aggdata <- aggregate(selectData$Study, by=list(selectData$REALM, 
                                               selectData$PROTECTED_AREA.y.1, 
                                               selectData$TAXA.x),
  FUN=length)
names(aggdata) <- c("Realm", "Protected_area", "Taxa", "Count")
print(aggdata)
```

##EVENESS and ABUNDANCE



How does evenness change relate to assemblage size change, coloured by taxa
```{r}
evAssT <- ggplot(selectData, aes( AbundanceAll, EvennessSim1, col = TAXA.x)) + 
	geom_point() + 
	theme_classic()+ 
	geom_vline(xintercept=c(0.05, -0.05), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	scale_colour_manual(values = c("DarkViolet", "Orange","LightBlue", "Green"))+
	labs(x = "Assembalge Size Change", y = "Evenness Change")
	ggMarginal(evAssT, type = "histogram", bins = 25)
```



How does evenness change relate to assemblage size change, coloured by whether or
not the assemblage is located within a protected area
```{r}
evAssPA <- ggplot(selectData, aes( AbundanceAll, EvennessSim1, col = PROTECTED_AREA.y.1)) + 
	geom_point() + 
	theme_classic() + 
	geom_vline(xintercept=c(0.05, -0.05), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	labs(x = "Assembalge Size Change", y = "Evenness Change")
	ggMarginal(evAssPA, type = "histogram", bins = 25)
```



How does evenness change relate to assemblage size change, point size related to 
length of study
```{r}
evAssL <- ggplot(selectData, aes( AbundanceAll, EvennessSim1, size = LengthYears.y)) + 
	geom_point() + 
	theme_classic()+ 
	geom_vline(xintercept=c(0.05, -0.05), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	labs(x = "Assembalge Size Change", y = "Evenness Change")
ggMarginal(evAssL, type = "histogram", bins = 25)
```



How does evenness change relate to assemblage size change, coloured by realm the 
assemblage is located in 
```{r}
evAssR <- ggplot(selectData, aes( AbundanceAll, EvennessSim1, col = REALM)) + 
	geom_point() + 
	theme_classic()+ 
	geom_vline(xintercept=c(0.05, -0.05), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	scale_colour_manual(values = c("Orange","LightBlue", "Green"))+
	labs(x = "Assembalge Size Change", y = "Evenness Change")
ggMarginal(evAssR, type = "histogram", bins = 25)
```

##EVENNESS and SPECIES RICHNESS

How does evenness change relate to species richness change, coloured by taxa

```{r}
evAbT <- ggplot(selectData, aes( SRslope, EvennessSim1, col = TAXA.x)) + 
	geom_point() + 
	theme_classic()+ 
	geom_vline(xintercept=c(0.03, -0.03), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	scale_colour_manual(values = c("DarkViolet", "Orange","LightBlue", "Green"))+
	labs(x = "Species Richness Change", y = "Evenness Change")
ggMarginal(evAbT, type = "histogram", bins = 25)
```



How does evenness change relate to species richness change, coloured by whether or
not the assemblage is located within a protected area

```{r}
evAbPA <- ggplot(selectData, aes( SRslope, EvennessSim1, col = PROTECTED_AREA.y.1)) + 
	geom_point() + 
	theme_classic()+ 
	geom_vline(xintercept=c(0.03, -0.03), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	labs(x = "Species Richness Change", y = "Evenness Change")
ggMarginal(evAbPA, type = "histogram", bins = 25)
```



How does evenness change relate to species richness change, coloured by teh realm
the assemblage is located in 
```{r}
evAbSR <- ggplot(selectData, aes(SRslope, EvennessSim1, col = REALM)) + 
	geom_point() + 
	theme_classic()+ 
	geom_vline(xintercept=c(0.03, -0.03), lty="dashed", col = "grey")+ 
	geom_hline(yintercept=c(0.0025, -0.0025), lty="dashed", col = "grey")+
	scale_colour_manual(values = c("Orange","LightBlue", "Green"))+
	labs(x = "Species Richness Change", y = "Evenness Change")
ggMarginal(evAbSR, type = "histogram", bins = 25)
```