---
title: "Abundance predators"
author: "Faith Jones"
date: "15 June 2017"
output: html_document
---
```{r, include=FALSE}
###Making a plot of how high pressure predation sites are when measured in 
#different levels of predation ABUNDANCE

setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(Cairo)


#preparing data 

fishDataAll <- read.csv("Fish data June 2017.csv")
head(fishDataAll)
str(fishDataAll)
levels(fishDataAll$species)
selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 

group12 <- c("Hoplias malabaricus", "Crenicichla frenata", "Gymnotus carapo", "Anablepsoides hartii",
			"Cichlasoma taenia", "Cichlasoma taenia", "Corynopoma riisei", "Roeboides dientonito", 
			"Agonostomus monticola", "Rhamdia quelen", "andinoacara pulcher", "Hemigrammus unilineatus", "Rivulus hartii")



#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

fishData$AbundanceAll <- fishData$numbercaught + fishData$numberseen

#calculating percentage biomass so i can incude years with 0 abundance

fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))
siteBiomass <- data.frame(TotBioSite)


#Narrow Predation data
#-------------------------
predatorsAbN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata")  %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(AbundanceAll))

predatorsAb <- data.frame(predatorsAbN)
predatorsAb$Year <- as.numeric(predatorsAb$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAb <- merge(siteBiomass, predatorsAb, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAb$AbundanceSum[is.na(FishDataTotAb$AbundanceSum)] <- 0

Narrow <- aggregate(FishDataTotAb$AbundanceSum, list(FishDataTotAb$site), mean, na.rm=TRUE)

orderNarrow <- Narrow[order(Narrow$x),]
head(orderNarrow)

orderNarrow$predOrderN <- c(1:nrow(orderNarrow))

#Middle predation 
#------------------------------

#there are no fsih of species Anablepsoides hartii

#sum abundance per site and year
predatorsAbM <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata" | 
		species == "Gymnotus carapo" | species == "Rivulus hartii")  %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(AbundanceAll))

predatorsAbMl <- data.frame(predatorsAbM)

predatorsAbMl$Year <- as.numeric(predatorsAbMl$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAbM <- merge(siteBiomass, predatorsAbMl, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAbM$AbundanceSum[is.na(FishDataTotAbM$AbundanceSum)] <- 0

Middle <- aggregate(FishDataTotAbM$AbundanceSum, list(FishDataTotAbM$site), mean, na.rm=TRUE)

orderMiddle<- Middle[order(Middle$x),]
head(orderMiddle)

orderMiddle$predOrderM <- c(1:nrow(orderMiddle))

#Wide predation 
#------------------------------

#there are no fsih of species Anablepsoides hartii

#sum abundance per site and year
predatorsAbW <- fishData %>% 
	filter(species %in% group12) %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(AbundanceAll))

predatorsAbWd <- data.frame(predatorsAbW)
predatorsAbWd$Year <- as.numeric(predatorsAbWd$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAbW <- merge(siteBiomass, predatorsAbWd, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAbW$AbundanceSum[is.na(FishDataTotAbW$AbundanceSum)] <- 0

Wide <- aggregate(FishDataTotAbW$AbundanceSum, list(FishDataTotAbW$site), mean, na.rm=TRUE)

orderWide <- Wide[order(Wide$x),]
head(orderWide)

orderWide$predOrderW <- c(1:nrow(orderWide))


##merge all data 
#-----------------------------

orderNM <- merge(orderNarrow, orderMiddle, by = "Group.1")
orderNMW <- merge(orderNM, orderWide, by = "Group.1")


```

Here is a plot showing the relationship between different levels of predation - 
Wide (group 2 and 3 species), Middle (group2) and Narrow(Hoplias and Crenicichla).
Sites are listed in decending order of mean predator pressure.

```{r}
Cairo(file="C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\bioTIME\\users\\Faith\\TriniFishPredation\\Adundance_tangle.png",
      type="png",
      units="in", 
      width=8,
      height=6,
      pointsize=12,
      dpi=100)

plotSetup <- data.frame(cbind(c(1:17),c(20:87))) # data to plot onto emptyp plot
plot(x = plotSetup$X2, y = plotSetup$X1, col="white", axes=FALSE , ylab = " ", xlab = "  ")

#plot arrows
arrows(x0 = 36, y0 = orderNMW$predOrderN, x1 = 48, y1 = orderNMW$predOrderM, length = 0.001)
arrows(x0 = 60, y0 = orderNMW$predOrderM, x1 = 72, y1 = orderNMW$predOrderW, length = 0.001)


#plotting text 
text(x = 54, y = orderNMW$predOrderM, labels = orderNMW$Group.1)
text(x = 30, y = orderNMW$predOrderN, labels = orderNMW$Group.1)
text(x = 78, y = orderNMW$predOrderW, labels = orderNMW$Group.1)

text(x = 30, y = 17, labels = "Narrow", cex = 1.5)
text(x = 54, y = 17, labels = "Intermediate", cex = 1.5)
text(x = 78, y = 17, labels = "Broad", cex = 1.5)

dev.off()

```

There is no obvious difference between narrow and middle defiinitions of predation
(spearman rank corrilation of 1). There is some difefference bewteenmiddle and wide
predation sites, but they are still strongly corrilated. 

```{r}
rownames(orderNMW) <- orderNMW$Group.1
corOrderNMW <- orderNMW[,c(2, 4, 6)] 
names(corOrderNMW ) <- c("Narrow", "Middle", "Wide")
cor(corOrderNMW, method="spearman")
```
