---
title: "Abundance change at different levels of predation"
author: "Faith Jones"
date: "16 June 2017"
output: html_document
---

Here i present the mixed model results for change in the abundance of predators. There
are three levels of predation: Wide (group 2 and 3 species), Middle (group2) and 
Narrow(Hoplias and Crenicichla). Only teh data from July/august is incuded so I 
could use the 2016 data. Also the survey eatken in 2010 is removed sure to sampling 
differences. 

```{r, include=FALSE}
setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lsmeans)
library(pbkrtest)

#preparing data 

fishDataAll <- read.csv("Fish data June 2017.csv")
head(fishDataAll)
str(fishDataAll)


selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 


#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

#calculate biomass
fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight

#calculating percentage abundance
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))

siteBiomass <- data.frame(TotBioSite)

```

#Narrow level predation definition 

```{r, include=FALSE}

predatorsAbN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(numbercaught))

predatorsAb <- data.frame(predatorsAbN)
predatorsAb$Year <- as.numeric(predatorsAb$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAb <- merge(siteBiomass, predatorsAb, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAb$AbundanceSum[is.na(FishDataTotAb$AbundanceSum)] <- 0
FishDataTotAb$Year <- as.numeric(FishDataTotAb$Year)


#add mean centered year for further analysis
meanYear <-summarise(FishDataTotAb, meanYear = mean(Year)) 
FishDataTotAb$meanYear <- meanYear[1,]
FishDataTotAb$meanCentredYear <-  FishDataTotAb$Year- FishDataTotAb$meanYear	

FishDataTotAb$logAb <- log2(FishDataTotAb$AbundanceSum + 1)

```

Here is the plot showing trendlines for each sites, using the linear model
part of ggplot

```{r}
NarrowAdundance <- ggplot(aes(x = Year, y = logAb, colour = site),data = FishDataTotAb)
NarrowAdundance + 
  geom_smooth(se = FALSE, method = "lm")+ 
	theme_classic()+
	labs(y = "Log Predator Abundance")

```

Here is the raw plot of predator biomass agaist time, and the plot of log2 biomass
against time. There is a lot of variability.
```{r}

AbPlot <- ggplot(FishDataTotAb, aes(x = Year, y = AbundanceSum, col = site, group = site))
AbPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance of Hoplias and Crenicichla", y = "Predator Abundance")

AbPlotLog <- ggplot(FishDataTotAb, aes(x = Year, y = logAb, col = site, group = site))
AbPlotLog + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance of Hoplias and Crenicichla", y = "Predator Abundance")


```

Next is the mixed model results and plot of slopes of the sites
```{r}

AbM1 <- lmer(logAb  ~ meanCentredYear + (meanCentredYear|site), data = FishDataTotAb)
summary(AbM1)

#plot slopes of sites
AbM1_Lines <- coef(AbM1)$site[c(1:2)] # get slope and intercept of each line in model 3
names(AbM1_Lines) <- c("intercept", "slope")

m3.plot <- ggplot(data = FishDataTotAb, aes(x = meanCentredYear, y = logAb)) # plot of model with varying intercept but not slope, but only main slope shown
m3.plot + geom_blank() + 
	theme_classic() +
	geom_abline(slope = AbM1_Lines$slope, intercept = AbM1_Lines$intercept) +
	labs(title="The slope of abunance change of Hoplias and Crenicichla", y = "Log Predator Abundance")

```

##Middle level predation definition 

```{r, include=FALSE}

predatorsAbNm <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata" | 
		species == "Gymnotus carapo"| species ==  "Rivulus hartii") %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(numbercaught + numberseen))

predatorsAbm <- data.frame(predatorsAbNm)
predatorsAbm$Year <- as.numeric(predatorsAbm$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAbm <- merge(siteBiomass, predatorsAbm, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAbm$AbundanceSum[is.na(FishDataTotAbm$AbundanceSum)] <- 0
FishDataTotAbm$Year <- as.numeric(FishDataTotAbm$Year)


#add mean centered year for further analysis
meanYear <-summarise(FishDataTotAbm, meanYear = mean(Year)) 
FishDataTotAbm$meanYear <- meanYear[1,]
FishDataTotAbm$meanCentredYear <-  FishDataTotAbm$Year - FishDataTotAbm$meanYear	

FishDataTotAbm$logAb <- log2(FishDataTotAbm$AbundanceSum + 1)

```
Here is the plot showing trendlines for each sites, using the linear model
part of ggplot

```{r}
MiddleAdundance <- ggplot(aes(x = Year, y = logAb, colour = site),data = FishDataTotAbm)
MiddleAdundance + 
  geom_smooth(se = FALSE, method = "lm")+ 
	theme_classic()+
	labs(y = "Log Predator Abundance")

```

Here is the raw plot of predator biomass agaist time, and the plot of log2 biomass
against time. There is a lot of variability.

```{r}

AbPlotm <- ggplot(FishDataTotAbm, aes(x = Year, y = AbundanceSum, col = site, group = site))
AbPlotm + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance group 2", y = "Predator Abundance")

AbPlotLogm <- ggplot(FishDataTotAbm, aes(x = Year, y = logAb, col = site, group = site))
AbPlotLogm + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance of  group 2", y = "Predator Abundance")


```

Next is the mixed model results and plot of slopes of the sites
```{r}

AbM1m <- lmer(logAb  ~ meanCentredYear + (meanCentredYear|site), data = FishDataTotAbm)
summary(AbM1m)

#plot slopes of sites
AbM1_Linesm <- coef(AbM1m)$site[c(1:2)] # get slope and intercept of each line in model 3
names(AbM1_Linesm) <- c("intercept", "slope")

m3.plot <- ggplot(data = FishDataTotAbm, aes(x = meanCentredYear, y = logAb)) # plot of model with varying intercept but not slope, but only main slope shown
m3.plot + geom_blank() + 
	theme_classic() +
	geom_abline(slope = AbM1_Linesm$slope, intercept = AbM1_Linesm$intercept) +
	labs(title="The slope of abunance change of group 2", y = "Log Predator Abundance")

```


##Wide level predation definition 

```{r, include=FALSE}

group12 <- c("Hoplias malabaricus", "Crenicichla frenata", "Gymnotus carapo", "Anablepsoides hartii",
			"Cichlasoma taenia", "Cichlasoma taenia", "Corynopoma riisei", "Roeboides dientonito", 
			"Agonostomus monticola", "Rhamdia quelen", "andinoacara pulcher", "Hemigrammus unilineatus", "Rivulus hartii")

predatorsAbNw <- fishData %>% 
	filter(species %in% group12)  %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(numbercaught))

predatorsAbw <- data.frame(predatorsAbNw)
predatorsAbw$Year <- as.numeric(predatorsAbw$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAbw <- merge(siteBiomass, predatorsAbw, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAbw$AbundanceSum[is.na(FishDataTotAbw$AbundanceSum)] <- 0
FishDataTotAbw$Year <- as.numeric(FishDataTotAbw$Year)

#add mean centered year for further analysis
meanYear <-summarise(FishDataTotAbw, meanYear = mean(Year)) 
FishDataTotAbw$meanYear <- meanYear[1,]
FishDataTotAbw$meanCentredYear <-  FishDataTotAbw$Year - FishDataTotAbw$meanYear	

FishDataTotAbw$logAb <- log2(FishDataTotAbw$AbundanceSum + 1)

```

Here is the plot showing trendlines for each sites, using the linear model
part of ggplot

```{r}
BroadAdundance <- ggplot(aes(x = Year, y = logAb, colour = site),data = FishDataTotAbw)
BroadAdundance + 
  geom_smooth(se = FALSE, method = "lm")+ 
	theme_classic()+
	labs(y = "Log Predator Abundance")

```
Here is the raw plot of predator biomass agaist time, and the plot of log2 biomass
against time. There is a lot of variability.
```{r}

AbPlotw <- ggplot(FishDataTotAbw, aes(x = Year, y = AbundanceSum, col = site, group = site))
AbPlotw + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance group 2 and 3", y = "Predator Abundance")

AbPlotLogw <- ggplot(FishDataTotAbw, aes(x = Year, y = logAb, col = site, group = site))
AbPlotLogw + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance of  group 2 and 3", y = "Predator Abundance")


```

Next is the mixed model results and plot of slopes of the sites
```{r}

AbM1w <- lmer(logAb  ~ meanCentredYear + (meanCentredYear|site), data = FishDataTotAbw)
summary(AbM1w)

#plot slopes of sites
AbM1_Linesw <- coef(AbM1w)$site[c(1:2)] # get slope and intercept of each line in model 3
names(AbM1_Linesw) <- c("intercept", "slope")

m3.plotw <- ggplot(data = FishDataTotAbw, aes(x = meanCentredYear, y = logAb)) # plot of model with varying intercept but not slope, but only main slope shown
m3.plot + geom_blank() + 
	theme_classic() +
	geom_abline(slope = AbM1_Linesw$slope, intercept = AbM1_Linesw$intercept) +
	labs(title="The slope of abunance change of group 2", y = "Log Predator Abundance")
```


