---
title: "Biomass Predators"
author: "Faith Jones"
date: "15 June 2017"
output: html_document
---
```{r, include=FALSE}
###Making a plot of how high pressure predation sites are when measured in 
#different levels of predation

setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(Cairo)


#preparing data 

fishDataAll <- read.csv("Fish data June 2017.csv")
head(fishDataAll)
str(fishDataAll)
levels(fishDataAll$species)
selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 


group12 <- c("Hoplias malabaricus", "Crenicichla frenata", "Gymnotus carapo", "Anablepsoides hartii",
			"Cichlasoma taenia", "Cichlasoma taenia", "Corynopoma riisei", "Roeboides dientonito", 
			"Agonostomus monticola", "Rhamdia quelen", "andinoacara pulcher", "Hemigrammus unilineatus", "Rivulus hartii")


#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

#calculate biomass
fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight


#calculating percentage biomass
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))

siteBiomass <- data.frame(TotBioSite)

#Narrow Predation data
#-------------------------

#sum biomass per site and year
predatorsBioN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBio <- data.frame(predatorsBioN)
predatorsBio$Year <- as.numeric(predatorsBio$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTot <- merge(siteBiomass, predatorsBio, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTot)
FishDataTot$BiomassSum[is.na(FishDataTot$BiomassSum)] <- 0

FishDataTot$PropMass <- FishDataTot$BiomassSum / FishDataTot$BiomasSiteSum

Narrow <- aggregate(FishDataTot$PropMass, list(FishDataTot$site), mean, na.rm=TRUE)

orderNarrow <- Narrow[order(Narrow$x),]
head(orderNarrow)

orderNarrow$predOrderN <- c(1:nrow(orderNarrow))

#Middle predation 
#------------------------------

#there are no fsih of species Anablepsoides hartii

#sum biomass per site and year
predatorsBioM <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata" | 
		species == "Gymnotus carapo" | species == "Rivulus hartii") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBioMl <- data.frame(predatorsBioM)

predatorsBioMl$Year <- as.numeric(predatorsBioMl$Year)

FishDataTotM <- merge(siteBiomass, predatorsBioM, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTotM)
FishDataTotM$BiomassSum[is.na(FishDataTotM$BiomassSum)] <- 0

FishDataTotM$PropMass <- FishDataTotM$BiomassSum / FishDataTotM$BiomasSiteSum


Middle <- aggregate(FishDataTotM$BiomassSum, list(FishDataTotM$site), mean, na.rm=TRUE)

orderMiddle<- Middle[order(Middle$x),]
head(orderMiddle)

orderMiddle$predOrderM <- c(1:nrow(orderMiddle))

#Wide predation 
#------------------------------

#there are no fsih of species Anablepsoides hartii

#sum biomass per site and year
predatorsBioW <- fishData %>% 
	filter(species %in% group12) %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBioWd <- data.frame(predatorsBioW)

predatorsBioWd$Year <- as.numeric(predatorsBioWd$Year)

FishDataTotW <- merge(siteBiomass, predatorsBioW, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTotW)
FishDataTotW$BiomassSum[is.na(FishDataTotW$BiomassSum)] <- 0

FishDataTotW$PropMass <- FishDataTotW$BiomassSum / FishDataTotW$BiomasSiteSum

Wide <- aggregate(FishDataTotW$BiomassSum, list(FishDataTotW$site), mean, na.rm=TRUE)

orderWide<- Wide[order(Wide$x),]
head(orderWide)

orderWide$predOrderW <- c(1:nrow(orderWide))


##merge all data 
#-----------------------------

orderNM <- merge(orderNarrow, orderMiddle, by = "Group.1")
orderNMW <- merge(orderNM, orderWide, by = "Group.1")

```

Here is a plot showing the relationship between different levels of predation - 
Wide (group 2 and 3 species), Middle (group2) and Narrow(Hoplias and Crenicichla).
Sites are listed in decending order of mean predator pressure.

```{r}

Cairo(file="C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\bioTIME\\users\\Faith\\TriniFishPredation\\Biomass_tangle.png",
      type="png",
      units="in", 
      width=9,
      height=4, 
      pointsize=12,
      dpi=100)

plotSetup <- data.frame(cbind(c(1:17),c(20:87))) # data to plot onto emptyp plot
plot(x = plotSetup$X2, y = plotSetup$X1, col="white", axes=FALSE , ylab = " ", xlab = "  ")

#plot arrows
arrows(x0 = 36, y0 = orderNMW$predOrderN, x1 = 48, y1 = orderNMW$predOrderM, length = 0.001)
arrows(x0 = 60, y0 = orderNMW$predOrderM, x1 = 72, y1 = orderNMW$predOrderW, length = 0.001)

#plotting text 
text(x = 54, y = orderNMW$predOrderM, labels = orderNMW$Group.1)
text(x = 30, y = orderNMW$predOrderN, labels = orderNMW$Group.1)
text(x = 78, y = orderNMW$predOrderW, labels = orderNMW$Group.1)

text(x = 30, y = 16.75, labels = "Narrow", cex = 1.25)
text(x = 54, y = 16.75, labels = "Intermediate", cex = 1.25)
text(x = 78, y = 16.75, labels = "Broad", cex = 1.25)

dev.off()

```

Each level of quantifying predation gives quite different estimates of which sites are high 
predation, but the three sites with very low predation are consistant. A corrilation
table using spearmans rank is listed below.The highest corrilation is between the
middle and the wide definitions of predation. 

```{r}
rownames(orderNMW) <- orderNMW$Group.1
corOrderNMW <- orderNMW[,c(2, 4, 6)] 
names(corOrderNMW ) <- c("Narrow", "Middle", "Wide")
cor(corOrderNMW, method="spearman")
```

Conclusion: It matters quite a bit what definition of "high predation site" you 
use when 
