---
title: "Biomass chane at different levels of predation"
author: "Faith Jones"
date: "16 June 2017"
output: html_document
---

Here i present the mixed model results for change in biomass of predators. There
are three levels of predation: Wide (group 2 and 3 species), Middle (group2) and 
Narrow(Hoplias and Crenicichla). Only teh data from July/august is incuded so I 
could use the 2016 data. Also the survey eatken in 2010 is removed sure to sampling 
differences. 

```{r, include=FALSE}
setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lsmeans)
library(pbkrtest)

#preparing data 

fishDataAll <- read.csv("Fish data June 2017.csv")
head(fishDataAll)
str(fishDataAll)


selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 


#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

#calculate biomass
fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight

#calculating percentage abundance
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))

siteBiomass <- data.frame(TotBioSite)

```

#Narrow level predation definition 

```{r, include=FALSE}
#Narrow

#sum biomass per site and year
predatorsBioN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBio <- data.frame(predatorsBioN)
predatorsBio$Year <- as.numeric(predatorsBio$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTot <- merge(siteBiomass, predatorsBio, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTot)
FishDataTot$BiomassSum[is.na(FishDataTot$BiomassSum)] <- 0

#add mean centered year for further analysis
meanYear <-summarise(FishDataTot, meanYear = mean(Year)) 
FishDataTot$meanYear <- meanYear[1,]
FishDataTot$meanCentredYear <-  FishDataTot$Year - FishDataTot$meanYear	

FishDataTot$LogBiomass <- log2(FishDataTot$BiomassSum + 1)

```

Here is the raw plot of predator biomass agaist time, and the plot of log2 biomass
against time. There is a lot of variability.
```{r}
BioPlot <- ggplot(FishDataTot, aes(x = Year, y = BiomassSum, col = site, group = site))
BioPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The biomass of Hoplias and Crenicichla", y = "Predator Biomass (kg)")

BioPlotLog <- ggplot(FishDataTot, aes(x = Year, y = LogBiomass, col = site, group = site))
BioPlotLog + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The log biomass of Hoplias and Crenicichla", y = "Log Predator Biomass (kg)")

```

Next is the mixed model results and plot of slopes of the sites
```{r}
bioMLog <- lmer(LogBiomass ~ meanCentredYear + (meanCentredYear|site), data = FishDataTot)
summary(bioMLog)

#plot slopes of sites
bioMLog_Lines <- coef(bioMLog )$site[c(1:2)] # get slope and intercept of each line in model 3
names(bioMLog_Lines) <- c("intercept", "slope")

m3.plotb <- ggplot(data = FishDataTot, aes(x = meanCentredYear, y = LogBiomass)) # plot of model with varying intercept but not slope, but only main slope shown
m3.plotb + geom_blank() + 
	theme_classic() + 
	geom_abline(slope =bioMLog_Lines$slope, intercept = bioMLog_Lines$intercept) +
	labs(title="The log biomass change of Hoplias and Crenicichla", y = "Log Predator Biomass (kg)")

```

#Middle level predation definition 

```{r, include=FALSE}
#Middle

#sum biomass per site and year
predatorsBioNm <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata" | 
		species == "Gymnotus carapo"| species == "Anablepsoides hartii") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBiom <- data.frame(predatorsBioNm)
predatorsBiom$Year <- as.numeric(predatorsBiom$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTotm <- merge(siteBiomass, predatorsBiom, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTotm)
FishDataTotm$BiomassSum[is.na(FishDataTotm$BiomassSum)] <- 0

#add mean centered year for further analysis
meanYear <-summarise(FishDataTotm, meanYear = mean(Year)) 
FishDataTotm$meanYear <- meanYear[1,]
FishDataTotm$meanCentredYear <-  FishDataTotm$Year - FishDataTotm$meanYear	

FishDataTotm$LogBiomass <- log2(FishDataTotm$BiomassSum + 1)
```

Here is the raw plot of predator biomass agaist time, and the plot of log2 biomass
against time, for middel predation. There is a lot of variability.
```{r}
BioPlotm <- ggplot(FishDataTotm, aes(x = Year, y = BiomassSum, col = site, group = site))
BioPlotm + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The biomass of group 2", y = "Predator Biomass (kg)")

BioPlotLogm <- ggplot(FishDataTotm, aes(x = Year, y = LogBiomass, col = site))
BioPlotLogm + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The log biomass of group 2", y = "Log Predator Biomass (kg)")

```


Mixed model results and plot of slopes of the sites (Middle)
```{r}
bioMLogm <- lmer(LogBiomass ~ meanCentredYear + (meanCentredYear|site), data = FishDataTotm)
summary(bioMLogm)

#plot slopes of sites
bioMLog_Linesm <- coef(bioMLogm)$site[c(1:2)] # get slope and intercept of each line in model 3
names(bioMLog_Linesm) <- c("intercept", "slope")

m3.plotbm <- ggplot(data = FishDataTotm, aes(x = meanCentredYear, y = LogBiomass)) # plot of model with varying intercept but not slope, but only main slope shown
m3.plotbm + geom_blank() + 
	theme_classic() + 
	geom_abline(slope =bioMLog_Linesm$slope, intercept = bioMLog_Linesm$intercept) +
	labs(title="The log biomass change of Group 2", y = "Log Predator Biomass (kg)")

```

#Wide level predation definition 

```{r, include=FALSE}
#Wide

group12 <- c("Hoplias malabaricus", "Crenicichla frenata", "Gymnotus carapo", "Anablepsoides hartii",
			"Cichlasoma taenia", "Cichlasoma taenia", "Corynopoma riisei", "Roeboides dientonito", 
			"Agonostomus monticola", "Rhamdia quelen", "andinoacara pulcher", "Hemigrammus unilineatus")

#sum biomass per site and year
predatorsBioNw <- fishData %>% 
	filter(species %in% group12) %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBiow <- data.frame(predatorsBioNw)
predatorsBiow$Year <- as.numeric(predatorsBiow$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTotw <- merge(siteBiomass, predatorsBiow, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTotw)
FishDataTotw$BiomassSum[is.na(FishDataTotw$BiomassSum)] <- 0

#add mean centered year for further analysis
meanYear <-summarise(FishDataTotw, meanYear = mean(Year)) 
FishDataTotw$meanYear <- meanYear[1,]
FishDataTotw$meanCentredYear <-  FishDataTotw$Year - FishDataTotw$meanYear	

FishDataTotw$LogBiomass <- log2(FishDataTotw$BiomassSum + 1)
```

Here is the raw plot of predator biomass agaist time, and the plot of log2 biomass
against time, for middel predation. There is a lot of variability.
```{r}
BioPlotw <- ggplot(FishDataTotw, aes(x = Year, y = BiomassSum, col = site, group = site))
BioPlotw + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The biomass of group 2 & 3", y = "Predator Biomass (kg)")

BioPlotLogw <- ggplot(FishDataTotw, aes(x = Year, y = LogBiomass, col = site))
BioPlotLogw + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The log biomass of group 2 & 3", y = "Log Predator Biomass (kg)")

```

Mixed model (wide)
```{r}
bioMLogw <- lmer(LogBiomass ~ meanCentredYear + (meanCentredYear|site), data = FishDataTotw)
summary(bioMLogw)

#plot slopes of sites
bioMLog_Linesw <- coef(bioMLogw)$site[c(1:2)] # get slope and intercept of each line in model 3
names(bioMLog_Linesw) <- c("intercept", "slope")

m3.plotbw <- ggplot(data = FishDataTotw, aes(x = meanCentredYear, y = LogBiomass)) # plot of model with varying intercept but not slope, but only main slope shown
m3.plotbm + geom_blank() + 
	theme_classic() + 
	geom_abline(slope =bioMLog_Linesw$slope, intercept = bioMLog_Linesw$intercept) +
	labs(title="The log biomass change of Group 2", y = "Log Predator Biomass (kg)")

```
