---
title: "Trindidad Fish Predation"
author: "Faith Jones"
date: "9 June 2017"
output: html_document
---
below is the initial analysis i have done to look at variability in predator mass
in streams. I have used only the 2010-2015 data as this is what i recieved from 
Faye. 

To simplify things I decided to use amean value for each study of each year. If 
we include the extra year i guess we coudl just use teh summer data? Otherwise
I would want to use the original seasonal data not a mean in teh final analysis
so we do not loose data.I also couldn't quite remember which were the main
two predators you wanted in the narrow definition of predators. I chose 
Rivulus and Crenicichla based on a literature search. 

```{r,results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)

setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")
fishData <- read.csv("FishData.csv")
```

##Biomass variability in predators initial analysis 
```{r,}

#sum biomass per site and year
predatorsBioN <- fishData %>% 
	filter(Species == "Rivulus hartii" | Species == "Crenicichla frenata") %>%
	separate(Date, c("Day", "Month", "Year"), sep = "/") %>%
	group_by(Site, Year) %>%
	summarise(BiomassSum = sum(weight))

predatorsBio <- data.frame(predatorsBioN)
predatorsBio$Year <- as.numeric(predatorsBio$Year )
head(predatorsBio)

BioPlot <- ggplot(predatorsBio, aes(x = Year, y = BiomassSum, col = Site, group = Site))
BioPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The biomass of Rivulus and Crenicichla", y = "Predator Biomass (kg)")

```

I then took a look at the coefficients of variation between sites and between years

```{r}

#variation between sites
cvSitesb <- predatorsBio %>%
  group_by(Site) %>%
  summarise(meanSite = mean(BiomassSum), sdSite = sd(BiomassSum))%>%
  mutate(cvSite = sdSite/meanSite)
cvSitesBio <- data.frame(cvSitesb)
print(cvSitesBio)

#variation between years
cvYearb <- predatorsBio %>%
  group_by(Year) %>%
  summarise(meanYear = mean(BiomassSum), sdYear = sd(BiomassSum))%>%
  mutate(cvSite = sdYear/meanYear)
cvYearBio <- data.frame(cvYearb)
print(cvYearBio)
```

Finally i put together a mixed model of predator biomass against year with 
site as a random variable, to see how much of the variation in the mean values 
coudl be explainedby site differences. I am not sure this is the right way to 
go though. You can see from teh results that sites is accounting for quite
a bit of variation, but other residual sources of variation are also 
important. I looked at teh standard deviation of the varience of the Random 
Effects to come to this conclusion. 

```{r}
#add mean centered year for further analysis
meanYear <-summarise(predatorsBio, meanYear = mean(Year)) 
predatorsBio$meanYear <- meanYear[1,]
predatorsBio$meanCentredYear <-  predatorsBio$Year - predatorsBio$meanYear

##try a mixed model for biomass
#-------------------------------
bioM1 <- lmer(BiomassSum ~ meanCentredYear + (1|Site), data = predatorsBio)
summary(bioM1)
```

##Abundance variability in predators initial analysis 

```{r}
predatorsAbN <- fishData %>% 
	filter(Species == "Rivulus hartii" | Species == "Crenicichla frenata") %>%
	separate(Date, c("Day", "Month", "Year"), sep = "/") %>%
	group_by(Site, Year) %>%
	summarise(AbundanceSum = sum(number.caught))

predatorsAb <- data.frame(predatorsAbN)
predatorsAb$Year <- as.numeric(predatorsAb$Year)

AbPlot <- ggplot(predatorsAb, aes(x = Year, y = AbundanceSum, col = Site, group = Site))
AbPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance of Rivulus and Crenicichla", y = "Predator Abundance")

```

Again, I took a look at the coefficients of variation between sites and over 
time.

```{r}
#add mean centered year for further analysis
meanYear <-summarise(predatorsAb, meanYear = mean(Year)) 
predatorsAb$meanYear <- meanYear[1,]
predatorsAb$meanCentredYear <-  predatorsAb$Year - predatorsAb$meanYear	

#variation between sites
cvSitesa <- predatorsAb %>%
  group_by(Site) %>%
  summarise(meanSite = mean(AbundanceSum), sdSite = sd(AbundanceSum))%>%
  mutate(cvSite = sdSite/meanSite)
cvSitesAb <- data.frame(cvSitesa)
print(cvSitesAb)

#variation between years
cvYeara <- predatorsAb %>%
  group_by(Year) %>%
  summarise(meanYear = mean(AbundanceSum), sdYear = sd(AbundanceSum))%>%
  mutate(cvYear = sdYear/meanYear)
cvYearAb <- data.frame(cvYeara)
print(cvYearAb)

```

Finnaly, I ran a mixed model of predatpor abundance against time with site as a 
random variable to see how much variation could be explained by be differences
in sites. In thsi model you can see that quite a lot of the variation from the 
mean can be attributed to site (standard deviation of 85 for site vs 45 for other
residual error). You can also see in the plot that different sites generraly remain 
either high or low predator pressure as we move through the years. 

```{r}

AbM1 <- lmer(AbundanceSum ~ meanCentredYear + (1|Site), data = predatorsAb)
summary(AbM1)


```

##Conclusions so far  

Biomass does not seem so be an ideal way of quantifying whether a site is high or 
low predation. A single site's predator biomass can vary as much over time as the 
difference between a high and low predator site in a singel year. So you may 
get different answers to wich sites are high predation from oen year to the next.

Abundance seems like a more reliable method of differentiating between high and 
low predation sites. The variability within a site was lower than between sites, 
so a site is likely to remain high or low predation regardless of the year you sample.

As I said, i am still consdering the most sensible of anylising the data properly
while keeping maximum information within the same analysis. I was pleased to see the 
mixed model results showed site to be far more influencial in explaining mean 
predator pressure in the abundance analysis than in the biomass analysis, but I 
do not think this is the best analysis for a paper unless we were discussing increases 
in predator pressure. Which we are not?

As an asside, why do you think predator biomass and abundance are so much lower 
in the first year for all sites?



