---
title: "Trindidad Fish Predation"
author: "Faith Jones"
date: "12 June 2017"
output: html_document
---
below is the next installment of the analysis to look at variability in predators
in streams. I have used only the 2010-2015 data as this is what i recieved from 
Faye. 

I decided to just use the summer data so i could incude the 2016 data.I also 
removed data from 2010


```{r,results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)


setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

fishDataAll <- read.csv("Fish data June 2017.csv")

selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 


#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

#calculate biomass
fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight


#calculating percentage abundance
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))

siteBiomass <- data.frame(TotBioSite)
```

##Biomass variability in predators initial analysis 
```{r,}

#sum biomass per site and year
predatorsBioN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBio <- data.frame(predatorsBioN)
predatorsBio$Year <- as.numeric(predatorsBio$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTot <- merge(siteBiomass, predatorsBio, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTot)
FishDataTot$BiomassSum[is.na(FishDataTot$BiomassSum)] <- 0

#add mean centered year for further analysis
meanYear <-summarise(FishDataTot, meanYear = mean(Year)) 
FishDataTot$meanYear <- meanYear[1,]
FishDataTot$meanCentredYear <-  FishDataTot$Year - FishDataTot$meanYear	

BioPlot <- ggplot(FishDataTot, aes(x = Year, y = BiomassSum, col = site, group = site))
BioPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The biomass of Hoplias and Crenicichla", y = "Predator Biomass (kg)")

```

I then took a look at the coefficients of variation between sites and between years

```{r}

#variation between sites
cvSitesb <- FishDataTot %>%
  group_by(site) %>%
  summarise(meanSite = mean(BiomassSum), sdSite = sd(BiomassSum))%>%
  mutate(cvSite = sdSite/meanSite)
cvSitesBio <- data.frame(cvSitesb)
print(cvSitesBio)

#variation between years
cvYearb <- FishDataTot %>%
  group_by(Year) %>%
  summarise(meanYear = mean(BiomassSum), sdYear = sd(BiomassSum))%>%
  mutate(cvSite = sdYear/meanYear)
cvYearBio <- data.frame(cvYearb)
print(cvYearBio)
```

Finally i put together a mixed model of predator biomass against year with 
site as a random variable, to see how much of the variation in the mean values 
coudl be explainedby site differences. I am not sure this is the right way to 
go though. You can see from teh results that sites is accounting for quite
a bit of variation, but other residual sources of variation far more important.
Site only explains 15% of the underlying variation in the model. 

```{r}

##try a mixed model for biomass
#-------------------------------
bioM1 <- lmer(BiomassSum ~ meanCentredYear + (1|site), data = FishDataTot)
summary(bioM1)
```
I then took a look at log biomass. Site then explains a lot more variation - nearly
50%. Still the sites are not consistantly high or low pressure. 
```{r}
FishDataTot$LogBio <- log2(FishDataTot$BiomassSum + 1)

##try a mixed model for biomass
#-------------------------------
bioM2 <- lmer(LogBio ~ meanCentredYear + (1|site), data = FishDataTot)
summary(bioM2)

BioPlotL <- ggplot(FishDataTot, aes(x = Year, y = LogBio, col = site, group = site))
BioPlotL + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The Log biomass of Hoplias and Crenicichla", y = "Log Predator Biomass (kg)")

```

##Abundance variability in predators initial analysis 

```{r}

predatorsAbN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(numbercaught))

predatorsAb <- data.frame(predatorsAbN)
predatorsAb$Year <- as.numeric(predatorsAb$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAb <- merge(siteBiomass, predatorsAb, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAb$AbundanceSum[is.na(FishDataTotAb$AbundanceSum)] <- 0

#add mean centered year for further analysis
meanYear <-summarise(FishDataTotAb, meanYear = mean(Year)) 
FishDataTotAb$meanYear <- meanYear[1,]
FishDataTotAb$meanCentredYear <-  FishDataTotAb$Year - FishDataTotAb$meanYear	

FishDataTotAb$logAb <- log2(FishDataTotAb$AbundanceSum + 1)

AbPlot <- ggplot(FishDataTotAb, aes(x = Year, y = AbundanceSum, col = site, group = site))
AbPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The abunance of Hoplias and Crenicichla", y = "Predator Abundance")

```

Again, I took a look at the coefficients of variation between sites and over 
time.

```{r}
#variation between sites
cvSitesa <- FishDataTotAb %>%
  group_by(site) %>%
  summarise(meanSite = mean(AbundanceSum, na.rm=TRUE), sdSite = sd(AbundanceSum, na.rm=TRUE))%>%
  mutate(cvSite = sdSite/meanSite)
cvSitesAb <- data.frame(cvSitesa)
print(cvSitesa)

#variation between years
cvYeara <- FishDataTotAb %>%
  group_by(Year) %>%
  summarise(meanYear = mean(AbundanceSum, na.rm=TRUE), sdYear = sd(AbundanceSum, na.rm=TRUE))%>%
  mutate(cvYear = sdYear/meanYear)
cvYearAb <- data.frame(cvYeara)
print(cvYeara)

```

I then ran a mixed model of predatpor abundance against time with site as a 
random variable to see how much variation could be explained by be differences
in sites. In thsi model you can see that quite a lot of the variation from the 
mean can be attributed to site (roughly 30%). Again sites do not remain either
high or low predation. 

```{r}

AbM1 <- lmer(AbundanceSum ~ meanCentredYear + (1|site), data = FishDataTotAb)
summary(AbM1)


```

##relative biomass

i have taken a qick look at proprtional biomass. This looks even more messy than
abundance or biomass. 
```{r}
FishDataTot$PropMass <- FishDataTot$BiomassSum / FishDataTot$BiomasSiteSum

PropBioPlot <- ggplot(FishDataTot, aes(x = Year, y = PropMass, col = site, group = site))
PropBioPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The proportional biomass of Hoplias and Crenicichla", y = "Proportional Predator Biomass (%)")


```

##Conclusions so far  

This is what i have so far:
- biomass/abundace/relative biomass all give different estimates of which sites
are high abundance
- there does not seem to be a clear seperation between high and low predation sites

I still havent decided on the best way to anylise the data. I looked at ANCOVA
but this requires the assumption that there is a linear relationship between the
covariate (time) and the response variable (predation level). I was thinking 
perhas an ANOVA or a regression. I did play with a regression, and gound that 
some sites were coming out as significantly different. I was wondering if you had
any other suggestions, as I mostly know how to look for trends, not weigh varience.



