---
title: "Richness predation level relationship"
author: "Faith Jones"
date: "9 August 2017"
output: html_document
---
Here i am plotting the relationship between the species richness and a sites agianst
the predation level at that site. Predation level is either quantified in terns 
of abundance or biomass.The definition of predation in this case is the narrow 
definition. 

```{r, include=FALSE}
setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(lsmeans)
library(pbkrtest)
library(Cairo)

#preparing data 

fishDataAll <- read.csv("Fish data June 2017.csv")
head(fishDataAll)
str(fishDataAll)


selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 


#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

#calculate biomass
fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight

#calculating percentage abundance
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))

siteBiomass <- data.frame(TotBioSite)

#calculate overall abundance
fishData$AbundanceAll <- fishData$numbercaught + fishData$numberseen

#calculating species richness for each site 

SR <- fishData %>%
  group_by(site)%>%
  summarise(SpeciesRichness = n_distinct(species))

Richness <- data.frame(SR)

fishDataSR <- merge(fishData, Richness, by = "site")


```

#getting biomass predation pressure 

```{r}

#sum biomass per site and year
predatorsBioN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBio <- data.frame(predatorsBioN)
predatorsBio$Year <- as.numeric(predatorsBio$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTot <- merge(siteBiomass, predatorsBio, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTot)
FishDataTot$BiomassSum[is.na(FishDataTot$BiomassSum)] <- 0

```
plotting biomass

```{r}

meanBiomass <- aggregate(FishDataTot$BiomassSum, list(FishDataTot$site), mean)
names(meanBiomass) <- c("site", "bioPred")

FishDataBio <- merge(meanBiomass, Richness, by = "site")


Cairo(file="C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\bioTIME\\users\\Faith\\TriniFishPredation\\BioSRyx.png",
      type="png",
      units="in", 
      width=8,
      height=6,
      pointsize=12,
      dpi=100)

biomassPlot <- ggplot(aes(x = SpeciesRichness,y = bioPred), data = FishDataBio)
biomassPlot + geom_point(size = 2) +
	labs(x = "Species Richness", y = "Predator biomass (kg)")+
  theme_classic()+
  theme(axis.title=element_text(size= 16), 
        axis.text=element_text(size= 15))

dev.off()

Cairo(file="C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\bioTIME\\users\\Faith\\TriniFishPredation\\BioSRxy.png",
      type="png",
      units="in", 
      width=8,
      height=6,
      pointsize=12,
      dpi=100)

biomassPlot <- ggplot(aes(y = SpeciesRichness, x = bioPred), data = FishDataBio)
biomassPlot + geom_point(size = 2) +
	labs(y = "Species Richness", x = "Predator biomass (kg)")+
  theme_classic()+
  theme(axis.title=element_text(size= 16), 
        axis.text=element_text(size= 15))

dev.off()


```

#biomass correlation 
```{r}
cor.test(FishDataBio$SpeciesRichness, FishDataBio$bioPred, method="spearman")
cor.test(FishDataBio$SpeciesRichness, FishDataBio$bioPred, method="pearson")

#no significant correlation 
```


#getting Abundance predation pressure 

```{r}

predatorsAbN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(AbundanceSum = sum(AbundanceAll))

predatorsAb <- data.frame(predatorsAbN)
predatorsAb$Year <- as.numeric(predatorsAb$Year)

#add the total site biomass data, just to i can identify sites with no predators 
FishDataTotAb <- merge(siteBiomass, predatorsAb, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
FishDataTotAb$AbundanceSum[is.na(FishDataTotAb$AbundanceSum)] <- 0
FishDataTotAb$Year <- as.numeric(FishDataTotAb$Year)


```

plotting Abundance

```{r}

meanAb <- aggregate(FishDataTotAb$AbundanceSum, list(FishDataTotAb$site), mean)
names(meanAb) <- c("site", "abPred")

FishDataAb <- merge(meanAb , Richness, by = "site")

Cairo(file="C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\bioTIME\\users\\Faith\\TriniFishPredation\\AbSRyx.png",
      type="png",
      units="in", 
      width=8,
      height=6,
      pointsize=12,
      dpi=100)

AbundancePlot <- ggplot(aes(x = SpeciesRichness,y = abPred), data = FishDataAb)
AbundancePlot + geom_point(size = 2) +
  theme_classic() +
	labs(x = "Species Richness", y = "Predator Abundance")+
  theme(axis.title=element_text(size= 16), 
        axis.text=element_text(size= 15))

dev.off()

Cairo(file="C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\bioTIME\\users\\Faith\\TriniFishPredation\\AbSRxy.png",
      type="png",
      units="in", 
      width=8,
      height=6,
      pointsize=12,
      dpi=100)

AbundancePlot <- ggplot(aes(y = SpeciesRichness, x = abPred), data = FishDataAb)
AbundancePlot + geom_point(size = 2) +
  theme_classic() +
	labs(y = "Species Richness", x = "Predator Abundance")+
  theme(axis.title=element_text(size= 16), 
        axis.text=element_text(size= 15))

dev.off()


```
#abundance correlation 
```{r}
cor.test(FishDataAb$SpeciesRichness, FishDataAb$abPred, method="spearman")
cor.test(FishDataAb$SpeciesRichness, FishDataAb$abPred, method="pearson")

#no significant correlation 

```

