---
title: "Predation level variability in Trini streams,biomass"
author: "Faith Jones"
date: "12 June 2017"
output: html_document
---
Streams in Trinidad have been used as natural experimental sites to study the 
effects of predation level on life characteristics if prey for many years. ususally
predation level will be defined as the biomass or number of very predatious fish
(crenicicula and hoplias). WHat is interesting is almost always, studies looking
at this sort of thing will define high and low predation sites and compare results.
A recent(ish) study looked at gradients of predation, and forn that generally the 
responses of prey to predation vary in proportion to teh predation level, so it 
doesnt make sense to just look at "high" and "low" predation sites. What has not 
been looked at much is how variable predation levels are within a site. Woudl I 
siet that is quantified as "high" in predation one year still en considered "high"
in predation if you looked again next year? Does the unit of measurement (abundance
vs biomass) make any difference? 

Our question is: is the variability in predation within a site greater than the 
variability between sites? (does this make sense?) Because if thsi is the case 
then it will eb difficult to reliably identify sites with high predation over time.

There are three ways to quantify predation pressure: biomass of predators, abundance 
of predators and % biomass of predators in cmparison to biomass of all fish. In this
document i will only discuss biomass. 

##preparing data
```{r,results='hide', message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)

setwd("C:\\Users\\faj\\Documents\\OneDrive for Business\\research\\FishPredation\\Data")

#preparing data 

fishDataAll <- read.csv("Fish data June 2017.csv")

selectMonths <- c("07", "08")# i am selecting out these months because then i can use the last timepoint 


#select data from only the July and Auguest, and remove first year 
SelectData <- fishDataAll %>% 
	separate(date, c("Day", "Month", "Year"), sep = "/") %>%
	filter(Year != 2010)%>%
	filter(Month %in% selectMonths)

fishData <- data.frame(SelectData)

#calculate biomass
fishData$TotBiomass <- (fishData$numbercaught + fishData$numberseen)* fishData$weight


#calculating percentage abundance
TotBioSite<- fishData %>% 
	group_by(site, Year) %>%
	summarise(BiomasSiteSum = sum(TotBiomass, na.rm=TRUE))

siteBiomass <- data.frame(TotBioSite)
```

##Biomass variability in predators initial analysis 

From the plot it doesnt look like thereare sites that are consistantly higher in 
predator biomass. It looks liek there are sites that have higher or lower varation 
though - some sites have consistently low predator variation whereas other sites 
have medium to high predator biomass but it depends a lot on what year you sample 

```{r,}

#sum biomass per site and year
predatorsBioN <- fishData %>% 
	filter(species == "Hoplias malabaricus" | species == "Crenicichla frenata") %>%
	group_by(site, Year) %>%
	summarise(BiomassSum = sum(TotBiomass))

predatorsBio <- data.frame(predatorsBioN)
predatorsBio$Year <- as.numeric(predatorsBio$Year)
siteBiomass$Year <- as.numeric(siteBiomass$Year)

FishDataTot <- merge(siteBiomass, predatorsBio, by = c("site" ,"Year"),all=TRUE) # this makes sure i 
#know what years had no predators recoreded in 
head(FishDataTot)
FishDataTot$BiomassSum[is.na(FishDataTot$BiomassSum)] <- 0

#add mean centered year for further analysis
meanYear <-summarise(FishDataTot, meanYear = mean(Year)) 
FishDataTot$meanYear <- meanYear[1,]
FishDataTot$meanCentredYear <-  FishDataTot$Year - FishDataTot$meanYear	

BioPlot <- ggplot(FishDataTot, aes(x = Year, y = BiomassSum, col = site, group = site))
BioPlot + geom_point() +
	geom_line() +
	theme_classic() +
	labs(title="The biomass of Hoplias and Crenicichla", y = "Predator Biomass (kg)")

```


Below I calculated the coeficients of variation. It looks to me that the variability 
within a site (over time) is similar to variability within a year(between sites)
```{r,}
##look at coefficients of variation 
#-------------------------------

#variation between sites
cvSitesb <- FishDataTot %>%
  group_by(site) %>%
  summarise(meanSite = mean(BiomassSum, na.rm=TRUE), sdSite = sd(BiomassSum, na.rm=TRUE))%>%
  mutate(cvSite = sdSite/meanSite)
cvSitesBio <- data.frame(cvSitesb)

#variation between years
cvYearb <- FishDataTot %>%
  group_by(Year) %>%
  summarise(meanYear = mean(BiomassSum, na.rm=TRUE), sdYear = sd(BiomassSum, na.rm=TRUE))%>%
  mutate(cvSite = sdYear/meanYear)
cvYearBio <- data.frame(cvYearb)

```


I then put together a mixed model of predator biomass against year with 
site as a random variable, to see how much of the variation in the mean values 
could be explainedby site differences. I am not sure this is the right way to 
go though. You can see from the results that "sites"" is accounting for some
variation, but other residual sources of variation are more influencial. 
I looked at the standard deviation of the varience of the Random Effects to come 
to this conclusion. 

```{r,}

##try a mixed model for biomass
#-------------------------------
bioM1 <- lmer(BiomassSum ~ meanCentredYear + (1|site), data = FishDataTot)
summary(bioM1)


```










